---
title: "Practical One"
format: html
---

This Quarto project (the rendered document itself, as well as the source code on GitHub).

# Question 1

Find all rows in **"airquality"** data set that have missing values.

**Code**

```{r}
rows_missing_val <-airquality[rowSums(is.na(airquality)) > 0, ]
```

**Answer**

```{r}
#| echo: false
rows_missing_val
```

# Question 2

Find mean, standard deviation, minimum, maximum for each of temperature and ozone level.

**Data Manipulation**

Extracts the rows in the data without missing values

```{r}
valid_rows <-airquality[rowSums(is.na(airquality)) == 0, ]
```

## Mean

**Code**

```{r}
mean_temp <- mean(valid_rows$Temp)
mean_ozone <- mean(valid_rows$Ozone)
```

**Answer:** Mean Temperature:

```{r}
#| echo: false 
mean_temp 
```

**Answer:** Mean Ozone Level:

```{r}
#| echo: false  
mean_ozone
```

## Standard Deviation

**Code**

```{r}
sd_temp <- sd(valid_rows$Temp)
sd_ozone <- sd(valid_rows$Ozone)
```

**Answer:** Standard Deviation of Temperature:

```{r}
#| echo: false
sd_temp
```

**Answer:** Standard Deviation of Ozone Level:

```{r}
#| echo: false   
sd_ozone
```

## Minimum

**Code**

```{r}
min_temp <- min(valid_rows$Temp) 
min_ozone <- min(valid_rows$Ozone)
```

**Answer:** Minimuim Temperature:

```{r}
#| echo: false
min_temp
```

**Answer:** Minimuim Ozone Level:

```{r}
#| echo: false   
min_ozone
```

## Maximum

**Code**

```{r}
max_temp <- max(valid_rows$Temp) 
max_ozone <- max(valid_rows$Ozone)
```

**Answer:** Maximuim Temperature:

```{r}
#| echo: false  
max_temp
```

**Answer:** Maximuim Ozone Level:

```{r}
#| echo: false   
max_ozone
```

# Question 3

## Calculate Beta estimate

Here, Y is the response variable, and X is the design matrix. The cars data (an R data set, also always available in R) contains two variables: speed and distance to stop. Fit a simple linear regression model to these data, i.e. find the beta estimates, using the equation above, and matrix calculations in R.

**Code**

```{r}
####### Create design matrix ####### 
Y <-cars$dist
intercept<- rep(1, length(Y))
X <- cbind(intercept,cars$speed)

#######  Function Calculates Beta hat and Std Error ####### 

calculate <- function(X, Y){
  
  beta_hat <- solve(t(X) %*% X) %*% t(X)%*% Y
  
  # Calculate Std Errors
  Sxx <- sum((X[,2] - mean(X[,2]))^2)
  Sxy <- sum((X[,2] - mean(X[,2])) * (Y - mean(Y)))
  
  predictions <- beta_hat[1]+beta_hat[2]*X[,2]
  residuals <- Y-predictions
  rse <- sqrt(sum(residuals^2)/(length(Y)-2))
  
  se_intercept<- rse*sqrt(1/length(Y)+mean(X[,2])^2/Sxx)
  se_speed <- rse/ sqrt( Sxx)
  
  # Display Std Errors
  std_error <- matrix(0, 2, 1) 
  std_error[1,1] <- se_intercept
  std_error[2,1] <- se_speed
  
  return( list("coefficients" = c(beta_hat[1], beta_hat[2]), 
               "std error" = c(se_intercept, se_speed)))
}


#######  Results ####### 
answer <-calculate(X,Y)

```

**Answer Beta estimate**

```{r}
#| echo: false
answer$coefficients
```

**Answer Standard Error**

```{r}
#| echo: false
answer$`std error`
```

# Question 4

## Confirm Beta estimate by fitting model

Check the same β estimates as when fitting the linear regression model using lm() in R.

**Code**

```{r}
#| echo: false
####### Create design matrix ####### 
Y <-cars$dist

intercept<- rep(1, length(Y))

X <- cbind(intercept,cars$speed)

#######  Function Calculates Beta hat and Std Error ####### 

calculate<-function(X, Y){
  
  # Solve for beta hat
  beta_hat <- solve(t(X) %*% X) %*% t(X)%*% Y
  
  # Calculate Std Errors
  Sxx <- sum((X[,2] - mean(X[,2]))^2)
  Sxy <- sum((X[,2] - mean(X[,2])) * (Y - mean(Y)))
  
  predictions <- beta_hat[1]+beta_hat[2]*X[,2]
  residuals <- Y-predictions
  rse <- sqrt(sum(residuals^2)/(length(Y)-2))
  
  se_intercept<- rse*sqrt(1/length(Y)+mean(X[,2])^2/Sxx)
  se_speed <- rse/ sqrt( Sxx)
  
  std_error <- matrix(0, 2, 1) 
  std_error[1,1] <- se_intercept
  std_error[2,1] <- se_speed
  
  return( list("coefficients" = c(beta_hat[1], beta_hat[2]), 
               "std error" = c(se_intercept, se_speed)))
}

```

```{r}
####### Calculate Beat values and Standard Error ####### 
answer <- calculate(X, Y)

####### Fit Model ####### 
model <- lm(dist ~ speed, data = cars)

# Print coefficients
print(answer$coefficients)   
print(model$coefficients) 


# Corrected comparison
message_1 <-ifelse (all(round(answer$coefficients,4) ==
                        round(model$coefficients,4)),
                        "The coefficients are the same!"
                        ,"The coefficients differ.")

message_2 <-ifelse (all(round(answer$`std error`,4) ==
                        round(summary(model)$coefficients[, "Std. Error"],4)),
                        "The standard errors are the same!",
                        "The tandard errors differ.")

```

**Answer if beta coefficients are correct**

```{r}
#| echo: false
message_1
```

**Answer if standard errors are correct**

```{r}
#| echo: false
message_2
```
