---
title: "Practical Four"
format: html
---

# Question 1

Display the flights dataset in an alternative format to simply printing it.

**Code**

```{r}
 #| echo: false  

  #################### Load packages ########################
  # Install tidyverse if not already installed
  if (!requireNamespace("tidyverse", quietly = TRUE)) {
    install.packages("tidyverse")
  }
  library(tidyverse)
  
  # Install and load nycflights13 for flight data
  if (!requireNamespace("nycflights13", quietly = TRUE)) {
    install.packages("nycflights13")
  }
  library(nycflights13)
  
  if (!requireNamespace("remotes", quietly = TRUE)) {
    install.packages("remotes")
  }
  remotes::install_github("SATVILab/UtilsDataRSV")
```

**Answer**
```{r}
  dplyr::glimpse(flights)

```

# Question 2

Rewrite the following code using dplyr

**Original Code**

```{r}
  flight1 <- flights[flights$month == 1, ]
carrier_vec <- unique(flight1$carrier)
carrier_dist_vec_mean <- numeric(length(carrier_vec))
carrier_dist_vec_sd <- numeric(length(carrier_vec))
for (i in seq_along(carrier_vec)) {
  carrier_dist_vec_mean[i] <- mean(
    flight1$distance[flight1$carrier == carrier_vec[i]]
   )
  carrier_dist_vec_sd[i] <- sd(
    flight1$distance[flight1$carrier == carrier_vec[i]]
  )
}
dist_tbl <- tibble(
  carrier = carrier_vec,
  mean_distance = carrier_dist_vec_mean,
  sd_distance = carrier_dist_vec_sd
)
dist_tbl[order(dist_tbl$mean_distance), ]
```

**Rewritten Code**

```{r}
  dist_tbl <- flights |> 
            filter(month == 1)|> 
            group_by(carrier)   |> 
            summarise(
                      avg_dist = mean(distance, na.rm = TRUE),
                      sd_dist = sd(distance, na.rm = TRUE))  |> 
            arrange(avg_dist)
```

**Answer**
```{r}
  dist_tbl
```

# Question 3

Explain why the standard deviation is NA for one carrier, and why it is 0 for others. Demonstrate your answer using code.

**Code**

```{r}
  q_3_1 <- flights |> 
         filter(carrier == "OO", month == 1, is.na(distance)) |>
         count()
```

**Answer**
```{r}
 q_3_1
```

There are no entries for carrier OO in month 1, therefore the standard deviation is NA

**Code**

```{r}
  q_3_2 <- flights |> 
         filter(carrier == "YV", month == 1) |>
         select(distance) |>
         print(n = Inf)
  
```

**Answer**
```{r}
 q_3_2
```
All the distance values for carrier YV and others with a standard deviation of 0 have the sam distance values for all entries in month 1

# Question 4
Using tidyr and dplyr where appropriate, construct a dataframe where the carriers are along the columns, and the rows are the average departure delay (dep_delay) flown by each carrier (carrier) in each month.

**Code**

```{r}
  q_4 <- flights |> 
         group_by(month, carrier) |> 
         summarise(avg_dep_delay = mean(dep_delay, na.rm = TRUE)) |>
         pivot_wider(  names_from = carrier, values_from = avg_dep_delay )
```

**Answer**
```{r}
  q_4
```

# Question 5
Calculate the proportion of flights that were delayed (dep_delay greater than 0) but arrived on or before time (arr_delay less than or equal to 0).

**Code**

```{r}
 q_5<- flights |> 
       filter(dep_delay > 0, arr_delay <= 0) |>  
       count() / nrow(flights)
```

**Answer**
```{r}
 
 q_5
```

# Question 6

Identify routes that more than one airline flies

**Code**

```{r}
 q_6_1 <- flights |>
          left_join(airlines, by = "carrier") |>
          group_by(origin, dest) |>
          filter(n_distinct(name) > 1) |>
          select(origin, dest) 
```

**Answer**
```{r}
  q_6_1
```

For each such route, calculate the average arrival delay for each airline (exclude NAs). Find the names of these airlines.

**Code**

```{r}
  q_6_2 <- flights |>
           left_join(airlines, by = "carrier") |>
           group_by(origin, dest, name) |>
           summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) |>
           select(origin, dest, name, avg_arr_delay)
```

**Answer**
```{r}
  
 q_6_2
```

For each such route, identify the airline with the worst and best average arrival delay.

**Code**

```{r}
  q_6_3_3 <- flights |>
             left_join(airlines, by = "carrier") |>
             group_by(origin, dest, name) |>
             summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) |>
             summarise(
                         best_airline = name[which.min(avg_arr_delay)],
                         worst_airline = name[which.max(avg_arr_delay)],
                         best_avg_arr_delay = min(avg_arr_delay),
                         worst_avg_arr_delay = max(avg_arr_delay),
                      )
```

**Answer**
```{r}

  q_6_3_3
```

Identify the route with the greatest difference between the best and worst performing airlines

**Code**

```{r}
  q_6_4 <- flights |>
           left_join(airlines, by = "carrier") |>
           group_by(origin, dest, name) |>
           summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE)) |>
           summarise(
                      best_avg_arr_delay = min(avg_arr_delay),
                      worst_avg_arr_delay = max(avg_arr_delay)
                    ) |>
           mutate(diff = worst_avg_arr_delay - best_avg_arr_delay) |>
           arrange(desc(diff)) |>
           filter(diff == max(diff))
  
```
**Answer**
```{r}
 q_6_4
```

Determine the reason for this difference

**Answer**
The route from JFK to ATL has the largest arrival delay on average

# Question 7

**Code**

```{r}
dataset <-structure(list(id = c("id_1", "id_2", "id_3", "id_4", "id_5", 
                      "id_6", "id_7", "id_8", "id_9", "id_10", "id_11", "id_12", "id_13", 
                      "id_14", "id_15", "id_16", "id_17", "id_18", "id_19", "id_20", 
                      "id_21", "id_22", "id_23", "id_24", "id_25", "id_26", "id_27", 
                      "id_28", "id_29", "id_30", "id_31", "id_32", "id_33", "id_34", 
                      "id_35", "id_36", "id_37", "id_38", "id_39", "id_40", "id_41", 
                      "id_42", "id_43", "id_44", "id_45", "id_46", "id_47", "id_48", 
                      "id_49", "id_50"), age = c(50L, 34L, 70L, 33L, 22L, 61L, 69L, 
                                                 73L, 62L, 56L, 71L, 33L, 73L, 44L, 45L, 46L, 24L, 70L, 46L, 76L, 
                                                 47L, 76L, 28L, 48L, 54L, 27L, 45L, 26L, 61L, 28L, 38L, 55L, 33L, 
                                                 36L, 62L, 58L, 72L, 31L, 34L, 51L, 61L, 64L, 26L, 28L, 60L, 29L, 
                                                 42L, 46L, 79L, 72L), gender = c("male", "male", "male", "female", 
                                                                                 "female", "male", "female", "male", "male", "female", "female", 
                                                                                 "male", "male", "female", "male", "male", "male", "male", "female", 
                                                                                 "male", "male", "male", "male", "female", "femal", "male", "female", 
                                                                                 "female", "female", "female", "male", "female", "female", "female", 
                                                                                 "male", "male", "female", "male", "female", "female", "male", 
                                                                                 "female", "female", "male", "male", "female", "male", "male", 
                                                                                 "male", "female"), height = c(174.4, 197.7, 174.1, 194.5, NA, 
                                                                                                               180.4, 170.5, 157.4, 196.8, 165.1, 153, 197.4, 186, 157.1, 177.5, 
                                                                                                               197.7, 179.3, 170.2, 182.4, NA, 165.4, 161, 168.5, 199.2, 157.7, 
                                                                                                               154.6, 157.1, 184.5, 181, 194.6, 183.6, 186.9, 176.1, 183, 191.1, 
                                                                                                               189.3, 199, 172, 165.6, 170.5, 150.5, 159.2, 192.1, 161.6, 162, 
                                                                                                               153.8, 162.3, 186.6, 192.4, 174.9), weight = c(69.4, 62.3, 55.6, 
                                                                                                                                                              69.5, 78.6, 60.8, 72.2, 60.9, 75.1, 67.7, 82.5, 68.7, 67.8, 76.7, 
                                                                                                                                                              87, 61.1, 70.6, 63.3, 81.5, 59.2, 93.2, 87.3, 83.4, 80.9, 68.6, 
                                                                                                                                                              76.5, 93.7, 79.1, 92, 65.6, 85.4, 63.3, 79.7, 74.1, 63.3, 78.2, 
                                                                                                                                                              95.7, 95.1, 63.7, 66.1, 99.3, 81, 96.9, 73.3, 70.3, 83, 57.6, 
                                                                                                                                                              78.6, 61.9, 98.1), blood_type = c("O", "A", "O", "O", "B", "AB", 
                                                                                                                                                                                                "O", "O", "O", "AB", "A", "O", "O", "O", "B", "A", "B", "AB", 
                                                                                                                                                                                                "O", "AB", "A", "AB", "O", "B", "A", "A", "B", "AB", "A", "B", 
                                                                                                                                                                                                "B", "A", "O", "O", "O", "B", "O", "A", "A", "B", "A", "O", "AB", 
                                                                                                                                                                                                "A", "A", "O", "O", "B", "A", "O"), disease_status = c("diseased", 
                                                                                                                                                                                                                                                       "healthy", "healthy", "healthy", "healthy", "healthy", "diseased", 
                                                                                                                                                                                                                                                       "healthy", "diseased", "Healthy", "diseased", "healthy", "diseased", 
                                                                                                                                                                                                                                                       "healthy", "diseased", "healthy", "healthy", "healthy", "healthy", 
                                                                                                                                                                                                                                                       "healthy", "healthy", "diseased", "healthy", "diseased", "healthy", 
                                                                                                                                                                                                                                                       "healthy", "healthy", "healthy", "diseased", "diseased", "healthy", 
                                                                                                                                                                                                                                                       "healthy", "healthy", "diseased", "diseased", "diseased", "healthy", 
                                                                                                                                                                                                                                                       "diseased", "healthy", "healthy", "healthy", "healthy", "healthy", 
                                                                                                                                                                                                                                                       "diseased", "diseased", "diseased", "healthy", "healthy", "diseased", 
                                                                                                                                                                                                                                                       "diseased"), cholesterol = c(228, 223, 213, 198, 166, 151, 195, 
                                                                                                                                                                                                                                                                                    199, 189, 196, 221, 156, 185, 230, 234, 174, 185, 236, 235, 180, 
                                                                                                                                                                                                                                                                                    165, 220, 160, 153, 250, 153, 184, 242, 212, 179, 224, 233, 181, 
                                                                                                                                                                                                                                                                                    199, 220, 214, 214, 248, 191, 162, 203, 173, 199, 187, 248, 189, 
                                                                                                                                                                                                                                                                                    173, 212, 164, 247), glucose = c(96, 78, 101, 119, 103, 91, 86, 
                                                                                                                                                                                                                                                                                                                     NA, 77, 80, 115, 85, 88, 109, NA, 71, 90, 94, 91, 87, 113, 93, 
                                                                                                                                                                                                                                                                                                                     97, 118, 109, 80, 85, 119, 99, 108, 89, 108, 97, 116, 79, 84, 
                                                                                                                                                                                                                                                                                                                     75, 81, 119, NA, 106, 109, 75, 82, 84, 75, 76, 120, 119, 77), 
               smoker = c("yes", "yes", "yes", "yes", "no", "yes", "no", 
                          "yes", "no", "no", "no", "no", "no", "yes", "no", "yes", 
                          "yes", "yes", "yes", "yes", "yes", "yes", "yes", "yes", "no", 
                          "no", "yes", "yes", "yes", "no", "no", "yes", "no", "yes", 
                          "no", "yes", "no", "yes", "yes", "yes", "no", "no", "yes", 
                          "no", "no", "no", "no", "no", "no", "yes"), exercise = c("occasional", 
                                                                                   "regular", "occasional", "regular", "none", "occasional", 
                                                                                   "regular", "none", "occasional", "none", "occasional", "none", 
                                                                                   "none", "regular", "occasional", "none", "regular", "regular", 
                                                                                   "none", "occasional", "none", "occasional", "occasional", 
                                                                                   "occasional", "regular", "occasional", "regular", "regular", 
                                                                                   "regular", "occasional", "occasional", "none", "none", "regular", 
                                                                                   "occasional", "occasional", "none", "none", "none", "none", 
                                                                                   "occasional", "regular", "regular", "none", "regular", "occasional", 
                                                                                   "occasional", "none", "occasional", "regular"), income = c(84820L, 
                                                                                                                                              81547L, 22588L, 72490L, 74533L, 25338L, 41469L, 57315L, 63629L, 
                                                                                                                                              88662L, 62615L, 56261L, 58499L, 82232L, 77584L, 77275L, 38468L, 
                                                                                                                                              54510L, 91326L, 78611L, 31402L, 29586L, 21441L, 58269L, 84173L, 
                                                                                                                                              88295L, 37940L, 43750L, 69750L, 92356L, 82518L, 91455L, 68866L, 
                                                                                                                                              51178L, 68275L, 27689L, 35418L, 81318L, 62405L, 86851L, 25654L, 
                                                                                                                                              47553L, 74474L, 51409L, 22607L, 55360L, 96351L, 21516L, 41927L, 
                                                                                                                                              55810L), education = c("master", "bachelor", "PhD", "master", 
                                                                                                                                                                     "bachelor", "highschool", "PhD", "highschool", "PhD", "PhD", 
                                                                                                                                                                     "bachelor", "highschool", "master", "bachelor", "PhD", "PhD", 
                                                                                                                                                                     "PhD", "bachelor", "master", "highschool", "PhD", "highschool", 
                                                                                                                                                                     "bachelor", "master", "highschool", "highschool", "master", 
                                                                                                                                                                     "master", "bachelor", "PhD", "highschool", "PhD", "master", 
                                                                                                                                                                     "master", "master", "PhD", "highschool", "master", "master", 
                                                                                                                                                                     "highschool", "bachelor", "highschool", "bachelor", "PhD", 
                                                                                                                                                                     "bachelor", "highschool", "master", "highschool", "bachelor", 
                                                                                                                                                                     "bachelor"), region = c("North", "South", "North", "West", 
                                                                                                                                                                                             "North", "West", "South", "South", "West", "South", "West", 
                                                                                                                                                                                             "South", "West", "East", "North", "West", "North", "North", 
                                                                                                                                                                                             "West", "North", "East", "West", "South", "North", "North", 
                                                                                                                                                                                             "East", "East", "North", "North", "West", "South", "West", 
                                                                                                                                                                                             "West", "East", "West", "North", "West", "North", "East", 
                                                                                                                                                                                             "North", "West", "South", "South", "East", "North", "West", 
                                                                                                                                                                                             "West", "East", "North", "East"), marital_status = c("divorced", 
                                                                                                                                                                                                                                                  "single", "divorced", "divorced", "divorced", "divorced", 
                                                                                                                                                                                                                                                  "divorced", "married", "divorced", "married", "divorced", 
                                                                                                                                                                                                                                                  "widowed", "married", "single", "widowed", "widowed", "single", 
                                                                                                                                                                                                                                                  "divorced", "widowed", "widowed", "single", "married", "single", 
                                                                                                                                                                                                                                                  "married", "widowed", "married", "single", "single", "widowed", 
                                                                                                                                                                                                                                                  "married", "widowed", "divorced", "single", "married", "single", 
                                                                                                                                                                                                                                                  "widowed", "widowed", "married", "widowed", "divorced", "married", 
                                                                                                                                                                                                                                                  "married", "divorced", "single", "married", "widowed", "divorced", 
                                                                                                                                                                                                                                                  "divorced", "single", "divorced")), row.names = c(NA, -50L
                                                                                                                                                                                                                                                  ), class = c("tbl_df", "tbl", "data.frame"))

UtilsDataRSV::view_cols(dataset)
```

**Answer**

Gender has a misspelled female entry, height has a missing value, disease_status has a misspelled healthy entry, glucose has a missing value


