---
title: "Question 2"
format: html
---

Find mean, standard deviation, minimum, maximum for each of temperature and ozone level.

**Data Manipulation**

Extracts the rows in the data without missing values

```{r}
valid_rows <-airquality[rowSums(is.na(airquality)) == 0, ]
```

## Mean

**Code**

```{r}
mean_temp <- mean(valid_rows$Temp)
mean_ozone <- mean(valid_rows$Ozone)
```

**Answer:** Mean Temperature:

```{r}
#| echo: false 
mean_temp 
```

**Answer:** Mean Ozone Level:

```{r}
#| echo: false  
mean_ozone
```

## Standard Deviation

**Code**

```{r}
sd_temp <- sd(valid_rows$Temp)
sd_ozone <- sd(valid_rows$Ozone)
```

**Answer:** Standard Deviation of Temperature:

```{r}
#| echo: false
sd_temp
```

**Answer:** Standard Deviation of Ozone Level:

```{r}
#| echo: false   
sd_ozone
```

## Minimum

**Code**

```{r}
min_temp <- min(valid_rows$Temp) 
min_ozone <- min(valid_rows$Ozone)
```

**Answer:** Minimuim Temperature:

```{r}
#| echo: false
min_temp
```

**Answer:** Minimuim Ozone Level:

```{r}
#| echo: false   
min_ozone
```

## Maximum

**Code**

```{r}
max_temp <- max(valid_rows$Temp) 
max_ozone <- max(valid_rows$Ozone)
```

**Answer:** Maximuim Temperature:

```{r}
#| echo: false  
max_temp
```

**Answer:** Maximuim Ozone Level:

```{r}
#| echo: false   
max_ozone
```
