---
title: "Question 3"
---

## Calculate Beta estimate

Here, Y is the response variable, and X is the design matrix. The cars data (an R data set, also always available in R) contains two variables: speed and distance to stop. Fit a simple linear regression model to these data, i.e. find the beta estimates, using the equation above, and matrix calculations in R.

**Code**

```{r}
####### Create design matrix ####### 
Y <-cars$dist
intercept<- rep(1, length(Y))
X <- cbind(intercept,cars$speed)

#######  Function Calculates Beta hat and Std Error ####### 

calculate <- function(X, Y){
  
  beta_hat <- solve(t(X) %*% X) %*% t(X)%*% Y
  
  # Calculate Std Errors
  Sxx <- sum((X[,2] - mean(X[,2]))^2)
  Sxy <- sum((X[,2] - mean(X[,2])) * (Y - mean(Y)))
  
  predictions <- beta_hat[1]+beta_hat[2]*X[,2]
  residuals <- Y-predictions
  rse <- sqrt(sum(residuals^2)/(length(Y)-2))
  
  se_intercept<- rse*sqrt(1/length(Y)+mean(X[,2])^2/Sxx)
  se_speed <- rse/ sqrt( Sxx)
  
  # Display Std Errors
  std_error <- matrix(0, 2, 1) 
  std_error[1,1] <- se_intercept
  std_error[2,1] <- se_speed
  
  return( list("coefficients" = c(beta_hat[1], beta_hat[2]), 
               "std error" = c(se_intercept, se_speed)))
}


#######  Results ####### 
answer <-calculate(X,Y)


```

**Answer Beta estimate**

```{r}
#| echo: false
answer$coefficients
```

**Answer Standard Error**
```{r}
#| echo: false
answer$`std error`
```
